---
title: "Booze and the Faith of America"
subtitle: "Alcohol and voting trends"
author: "Greyfriar's Bobby <br> Lucas Amar, Pablo Ortuno, Akshaj Pilla, Ata Tuna"
institute: "Merchant logo The university of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r load-packages, include = FALSE, message = FALSE}
# Add any additional packages you need to this chunk

install.packages("huxtable")

library(tidyverse)
library(tidymodels)
library(knitr)
library(usmap)
library(xaringanthemer)
library(showtext)
library(huxtable) 

 source('/cloud/project/extra/Model/functions/predict_election_win.R')
 source('/cloud/project/extra/Model/functions/predict_election_plot.R')
 source('/cloud/project/extra/Model/functions/wrong_preds.R')
 source('/cloud/project/extra/Model/functions/wrong_preds_spirit.R')
 source('/cloud/project/extra/Model/functions/predict_election_plot_spirit.R')
 source('/cloud/project/extra/Model/functions/predict_election_win_spirit.R')

```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
style_mono_accent(
  base_color = "#43418A",
  header_font_google = google_font("Padauk"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


```{r load-data, include=FALSE}
# Load your data here
voting_data <- read_csv("/cloud/project/data/clean_presentation_data/voting_data.csv")
election_pred <- read_csv("/cloud/project/data/clean_presentation_data/election_pred.csv")
model_data_train <- read_csv("/cloud/project/data/clean_presentation_data/model_data_train.csv")
model_data_test <- read_csv("/cloud/project/data/clean_presentation_data/model_data_test.csv")
electoral_votes <- read_csv("/cloud/project/data/clean_presentation_data/electoral_votes.csv")
predicted_elections <- read_csv("/cloud/project/data/clean_presentation_data/predicted_elections.csv")
predicted_elections_improv <- read_csv("/cloud/project/data/clean_presentation_data/predicted_elections_improv.csv")

model_data_train["office"] <- lapply(model_data_train["office"] , factor)
```
---

class: center, middle 

##Booze and the Faith of America

### Can alcohol consumption predict U.S. general election results?

Greyfriar's Bobby: Pablo Ortuno, Akshaj Pilla, Ata Tuna,  Lucas Amar

The University of Edinburgh

03/12/2020

---
# Data

### Building a data frame


```{r data_2, echo = FALSE}
model_data_train %>%
  print (n = 6)
```

---


# Model

```{r preliminary_model_2016}
model_election <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(office ~ ., data = model_data_train, family = "binomial")
```

```{r model_2, echo = FALSE}
tidy(model_election) %>%
  print(10)
```

---

## Model's results

```{r preliminary_2016_prediction, echo = FALSE, message=FALSE, fig.height="80%"}

predict_election_plot(2016)

```

---

## Interpreting results

```{r interpretation_1, echo = FALSE, message = FALSE}
wrong_preds(2016)
```
  
Our model correctly predicted 34 states and had a 68% accuracy

```{r intepretation_2, echo = FALSE, message = FALSE}
electoral_votes %>%
  filter(state != "District Of Columbia") %>%
  mutate(predicted = election_pred$predicted) %>%
  group_by(predicted) %>%
  summarise(total_votes = sum(votes))

```

In 2016, the republican party won the election with 306 electoral votes. 
The democrat party only got 232 electoral votes
---

## Improvements

```{r improvement_1, echo = FALSE}

model_election_improv <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(office ~ . - beer_per_capita - wine_per_capita, data = model_data_train, family = "binomial")

election_pred_improv <- predict(model_election_improv, model_data_test, type = "prob") %>%
  bind_cols(model_data_test) %>%
  mutate(predicted = ifelse(.pred_democrat > .pred_republican, "democrat", "republican"))
```


```{r improv_12, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}
predict_election_plot_spirit(2016)
```


---

## Improvements

```{r improv_2, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

wrong_preds_spirit(2016)

```

Our model predicted correctly 40 states, and yields an 80 % accuracy

```{r improv_3, echo = FALSE, message = FALSE}
electoral_votes %>%
  filter(state != "District Of Columbia") %>%
  mutate(predicted = election_pred_improv$predicted) %>%
  group_by(predicted) %>%
  summarise(total_votes = sum(votes))
```

It still fails to predict correctly the overall result
---
## Accuracy

First model
```{r accuracy_1, echo = FALSE}
kable(predicted_elections)

```

Improved model
``` {r accuracy_2, echo = FALSE}
kable(predicted_elections_improv)
```


---
## Discussion


```{r discussion_1, message = FALSE, echo = FALSE, out.height="50%"}
model_data_train %>%
  group_by(year) %>%
  summarise(drinking_avg = sum(beer_per_capita,
                               wine_per_capita, 
                               spirit_per_capita) / 50) %>%
  ggplot(aes(x = year,
             y = drinking_avg)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Average yearly amount of alcohol drank per person in the U.S",
       subtitle = "per year",
       x = "Year",
       y = "Ethanol gallons")
  
```
---
## Discussion

```{r discussion_4, echo = FALSE,out.width="80%"}
plot_usmap(data = election_pred, 
           regions = "states", 
           values = "beer_per_capita", 
           color = "black") +
  labs( title = "Amount of beer drank per person",
        subtitle = "by state, in 2016") +
  theme( legend.position = "right") +
  scale_fill_continuous(low = "white", 
                        high = "black", 
                        name = "Ethanol gallons", 
                        label = scales::comma)
```

---
## Discussion
```{r discusion_2, echo = FALSE, out.width="80%"}

election_pred %>%
  ggplot(aes(x = state, y = beer_per_capita)) +
  geom_point() +
  theme_minimal() +
  theme(axis.text.x = element_blank() ) +
  facet_wrap(~office) +
  labs( title = "Average beer drank per person", 
        subtitle = "by states in 2016", 
        x = "", 
        y = "Beer gallons")
```
---
## Discussion

Mean amount of ethanol drank per person on 2016, by state's political affiliation

```{r discussion_3, message = FALSE, echo = FALSE}
election_pred %>%
  group_by(office) %>%
  summarise(beer_mean = mean(beer_per_capita), wine_mean = mean(wine_per_capita), spirit_mean = mean(spirit_per_capita))
```
---


## Conclusion

- We achieved a 68% success rate (When using all types of alcohol together)

  - Through our analyses we saw that total alcohol consumption doesn't really vary from state to state
  
    - Makes it complicated to find a relationship between political affiliation and drinking

  - Through further investigation we therefore looked at the consumption of the other alcohols we had available
      
      - By using spirit only we got our success rate up to 80%
        

---


## Conclusion

-Overall:
  - Not able to establish a strong correlation, and even if we were... correlation doesn't imply causation
  
  - Know how to make a logistic regression and manipulate data much more efficiently and with less errors than before
  
  - Considering a possible relationship between alcohol consumption and political affiliation arises questions, will a correlation arise when looking at:
  
    - Alcohol brand?
    
    - Alcoholism?
    
    - Liver failure due to alcohol?
    
    - Etc.

---
