---
title: "Untitled"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(here)
library(stringr)
library(openintro)
library(usmap)
```

```{r loading_data}
# Voting data
voter_data <- read_rds("/cloud/project/data/1976-2016-president.RData")
write_rds(voter_data, "/cloud/project/data/voter_data.rds")
```


### CLEANING VOTING DATA

Preliminary cleaning
```{r clean_voter_data}
voter_data %>%
  filter (party %in% c("democrat", "republican")) %>% 
  select(year, state, party, candidatevotes) %>% 
  pivot_wider(names_from = party, values_from = candidatevotes) %>%
  rename(democrat_votes = democrat, republican_votes = republican)
```
This outputs votes as a list. 
This must be because two different rows have a shared value.
Lets examine and find where the problem is
```{r examining_voting_issues}
voter_data %>%
  filter (party %in% c("democrat", "republican")) %>% 
  select(year, state, party, candidatevotes) %>% 
  pivot_wider(names_from = party, values_from = candidatevotes, values_fn = length) %>%
  #Adding the values_fn = length argument is going to tell us wich entries of the list have a vector (its length will not be = 1)
  filter(
    democrat != 1 | republican != 1
  )
```
So the issues arise in 2004 Maryland, where two rows contain "democrat", 2016 Arizona, where two rows contain "democrat", and 2016 Maryland, where there are two different rows for each party.
Lets see exactly where the error arises from
```{r voting_issue_1}
voter_data %>%
  select(year, state, candidate, party, candidatevotes) %>%
  filter(year == 2004, state == "Maryland", party == "democrat")
```
The problem here is two different candidates. Fixed by erasing row 2545

```{r voting_issue_2}
voter_data %>%
  select(year, state, candidate, party, candidatevotes) %>%
  filter(year == 2016, state == "Arizona", party == "democrat")
```
The problem here is, again, two different candidates for the same part
```{r voting_issue_3}
voter_data %>%
  select(year, state, candidate, party, candidatevotes) %>%
  filter(year == 2016, state == "Maryland", party == "democrat" | party == "republican")

```

So we'll solve this issues by erasing rows 2415, 3415, and adding the corresponding values for rows 3565 and 3544, and 3537 and 3543.
Since we know exactly what values we're dealing with we'll just rewrite the values and delete two rows

```{r solved_voting_issues}

voter_data[(3536),] <- voter_data[(3536),] %>%
    mutate(candidatevotes = 1677928 + 78)

voter_data[(3537),] <- voter_data[(3537),] %>%
    mutate(candidatevotes = 943169 + 259)
  
voter_data <- voter_data[-c(2545,3415,3543,3544),]
```

Now we can correctly apply our cleaning and add a "winner" variable

```{r final_voting_cleaning}
clean_votes <- voter_data %>%
  filter (party %in% c("democrat", "republican")) %>% 
  select(year, state, party, candidatevotes) %>% 
  pivot_wider(names_from = party, values_from = candidatevotes) %>%
  rename(democrat_votes = democrat, republican_votes = republican) %>%
  mutate(winner = ifelse(democrat_votes > republican_votes, "democrat", "republican")) # Note there are no entries where the number of votes are equal 

clean_votes
```

As expected, class variables are as desired


