---
title: "Project"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r libraries, echo = FALSE,message=FALSE}
library(tidyverse)
library(tidymodels)
library(usmap)
library(ggpubr)

```


# 1. Data

```{r load_data, echo = FALSE, message=FALSE}

electoral_votes <- read_csv("/cloud/project/final/data/electoral_votes.csv")
consumption_states <- read_csv("/cloud/project/final/data/model_consumption.csv")
voting_data <- read_csv("/cloud/project/final/data/voting_data.csv")

```

## Define U.S. geographical regions and colors

```{r regions}

northeast <- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "New Jersey", "New York", "Pennsylvania", "Rhode Island", "Vermont", "Delaware", "Maryland")
southeast <- c("Alabama", "Arkansas", "Florida", "Georgia", "Kentucky", "Louisiana", "Mississippi", "North Carolina", "South Carolina", "Tennessee", "Virginia", "West Virginia")
midwest <- c("Illinois", "Indiana", "Iowa", "Kansas", "Michigan", "Minnesota", "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota", "Wisconsin")
southwest <- c("Arizona", "New Mexico", "Oklahoma", "Texas")
west <- c("Alaska", "California", "Colorado", "Hawaii", "Idaho", "Montana", "Nevada", "Oregon", "Utah", "Washington", "Wyoming")

partisan_colors <- c("democrat" = "blue3", "republican" = "red3")

```


## Merge consumption and voting data

```{r cleaning, message=FALSE}

consumption_states <- full_join(consumption_states, voting_data) %>%
  filter(!(state %in% c("District Of Columbia", "District of Columbia", "Us Total", "South Region", "Northeast Region", "Midwest Region", "West Region"))) %>%
  arrange(state, year)

# Split

split_1 <- consumption_states %>%
  filter(year == 2016)

split_2 <- consumption_states %>%
  filter(year != 2016)

for (n in 1:nrow(split_2)) {
  if ((n-1)%%4 == 0) {
  split_2$office[n] -> split_2$office[n+3]
  split_2$office[n] -> split_2$office[n+2]
  split_2$office[n] -> split_2$office[n+1]
  }
}

consumption_states <- full_join(split_1, split_2) %>%
  filter(year != 1976) %>%
  select(!democrat_votes & !republican_votes)

consumption_states["office"] <- lapply(consumption_states["office"] , factor)

consumption_states <- consumption_states %>%
  mutate(region = case_when(
    state %in% northeast ~ "northeast",
    state %in% southeast ~ "southeast",
    state %in% midwest ~ "midwest",
    state %in% southwest ~ "southwest",
    state %in% west ~ "west"
  ))

```

# 2. Trends for alcohol consumption

```{r consumption_evolution ,message = FALSE, echo = FALSE, fig.width=8}

consumption_states %>%
  group_by(year) %>%
  summarise(
    total = mean(ethanol_all_drinks_gallons_per_capita),
    beer = mean(beer_per_capita),
    wine = mean(wine_per_capita),
    spirit = mean(spirit_per_capita)) %>%
    pivot_longer(c(total, beer, wine, spirit), "alcohols") %>%
  mutate(
    alcohols = factor(alcohols),
    alcohols = fct_relevel(alcohols, "total", "beer", "spirit", "wine")) %>%
  ggplot(aes(x = year, y = value, color = alcohols)) +
  geom_line()+
  labs( title = "Average ethanol consumption per capita",
        subtitle = "by type of alcohol consumed",
        x = "Year",
        y = "Ethanol gallons",
        colour = "Type of alcohol") +
  ylim(0.25, 2.9) +
  theme_minimal() +
  scale_color_viridis_d()

```

## 3. Idk some other visualisation

## 4. Modelling

### First model, fitting all data

```{r split_data}
# Drop out co linear variables
consumption_states_model <- consumption_states %>%
  select(state, beer_per_capita, wine_per_capita, spirit_per_capita, office)

# Split data into training and testing
set.seed(7637)
consumption_split <- initial_split(consumption_states_model)
consumption_train <- training(consumption_split)
consumption_test <- testing(consumption_split)

```


```{r model, echo = FALSE, message = FALSE }

# Create a recipe

election_rec <- recipe(office ~ ., data = consumption_test) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Define engine

election_spec <- logistic_reg() %>%
  set_engine("glm")

# Work-flow

election_wflow <- workflow() %>%
  add_model(election_spec) %>%
  add_recipe(election_rec)

```

```{r cross_validation, echo = FALSE, message = FALSE }

set.seed(2342)
election_folds <- vfold_cv(consumption_train, v = 5)

election_fit_rs <- election_wflow %>%
  fit_resamples(election_folds)

collect_metrics(election_fit_rs, summarize = FALSE)
collect_metrics(election_fit_rs)

```

```{r roc_curve_1, echo = FALSE, message = FALSE }

election_fit <- election_wflow %>%
  fit(consumption_train)

election_pred <- predict(election_fit, consumption_test, type = "prob") %>%
  bind_cols(consumption_test %>% select(office))

```

### ROC Curve and number of correct predictions

```{r roc_curve_plot}
election_pred %>%
  roc_curve(
    truth = office,
    .pred_democrat) %>%
  autoplot()

election_pred %>%
  mutate(
    prediction = ifelse(.pred_democrat > .pred_republican, "democrat", "republican")
  ) %>%
  count(prediction == office)
```

### Predicting 2012 election

Map of states predicted correctly. The figure below indicates the number of states predicted correctly

```{r predicting_2012, echo = FALSE, message = FALSE, fig.width = 10}

test_2012 <- consumption_states %>% filter(year == 2012) %>% select(state, beer_per_capita, wine_per_capita, spirit_per_capita, office)
train_2012 <- consumption_states %>% filter(year < 2012) %>% select(state, beer_per_capita, wine_per_capita, spirit_per_capita, office)

election_rec_2012 <- recipe(office ~ ., data = test_2012) %>%
  step_dummy(all_nominal(), -all_outcomes())

election_2012_wflow <- workflow() %>%
  add_model(election_spec) %>%
  add_recipe(election_rec_2012)

election_2012_fit <- election_2012_wflow %>%
  fit(train_2012)

election_2012_pred <- predict(election_2012_fit, test_2012, type = "prob") %>%
  bind_cols(test_2012 %>% select(office, state))

plot_2012 <- election_2012_pred %>%
  mutate( prediction = ifelse(.pred_democrat > .pred_republican, "democrat", "republican")) %>%
  filter(office == prediction)

plot_usmap(data = plot_2012, regions = "states", values = "office", color = "black") +
      labs( title = "States Predicted Correctly",
          subtitle = "by winning party",
          caption = "States coloured white were predicted incorrectly",
          fill = "office") +
    theme(
      legend.position = "right") +
    scale_fill_manual(values = partisan_colors)

plot_2012 %>%
  count(office == prediction)

```

Model's general election result

```{r 2012_election_result, echo = FALSE}

full_join(election_2012_pred, electoral_votes) %>%
  mutate( prediction = ifelse(.pred_democrat > .pred_republican, "democrat", "republican")) %>%
  filter(state != "District Of Columbia") %>%
  select(state, prediction, votes) %>%
  group_by(prediction) %>%
  summarise(total_votes = sum(votes))

```

The actual results were:
Democrat - 332
Republican - 206

### Assessing accuracy

```{r 2012_roc_curve}

election_2012_pred %>%
  roc_curve(
    truth = office,
    .pred_democrat) %>%
  autoplot()

election_2012_pred %>%
  roc_auc(
    truth = office,
    .pred_democrat)

```


### Second model, only using spirit
