---
title: "Atarandom-code"
author: "Greyfriars Bobby"
output: github_document
editor_options: 
  chunk_output_type: console
---
### Libraries
```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(here)
library(stringr)
library(openintro)
library(usmap)
```

### Load Data
```{r load-data, message = FALSE, error = FALSE}
# Beer production by state
beer_states <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_states.csv')
write_csv(beer_states, "/cloud/project/data/beer_states.csv")

# Voting data
voter_data <- read_csv("/cloud/project/data/clean_presentation_data/voting_data.csv")

#Alcohol consumption per capita
consumption_per_capita <- read_csv("/cloud/project/data/apparent_per_capita_alcohol_consumption_1977_2018.txt")
write_csv(consumption_per_capita, "/cloud/project/data/apparent_per_capita_alcohol_consumption_1977_2018.txt")
```

### Code
```{r standardise-names-of-data-variables}
#capitalise state first letter on consumption_per_data
consumption_per_capita <- consumption_per_capita %>%
  mutate(state = str_to_title (state))

#change state abbreviation to name on beer_states

beer_states <- beer_states %>%
  mutate(state = abbr2state (state))

```

```{r beer_state dataframes}

beer_states_2015_OP <- beer_states %>% 
    filter(year == 2015, type == "On Premises") 

beer_states_2015_BC <- beer_states %>% 
    filter(year == 2015, type == "Bottles and Cans") 

beer_states_2015_KB <- beer_states %>% 
    filter(year == 2015, type == "Kegs and Barrels") 

```


```{r beer_production_by_on_premises}

plot_usmap(data = beer_states_2015_OP, values = "barrels", regions = "state", color = "black") +
  labs(title = "Production of beer on premises",
       subtitle = "across the United States, by state") + 
   scale_fill_continuous(
    low = "white", high = "red", name = "by Beer Production", label = scales::comma
  ) + theme(legend.position = "right")

```

```{r beer_production_by_bottles_and_cans}

plot_usmap(data = beer_states_2015_BC, values = "barrels", regions = "state", color = "black") +
  labs(title = "Production of bottled and canned beer",
       subtitle = "across the United States, by state") + 
   scale_fill_continuous(
    low = "white", high = "red", name = "by Beer Production", label = scales::comma
  ) + theme(legend.position = "right")

```

```{r beer_production_by_kegs_and_barrels}

plot_usmap(data = beer_states_2015_KB, values = "barrels", regions = "state", color = "black") +     
  labs(title = "Production of beer kegs and barrels",
       subtitle = "across the United States, by state") + 
  scale_fill_continuous(
    low = "white", high = "red", name = "by Beer Production", label = scales::comma
  ) + 
  theme(legend.position = "right")

```


### Data wrangling

This chunk cleans the voting data frame into a more useful one.

```{r clean_voter_data}

# Manually merge some conflicting cells
# In some states there were two entries for the same candidate, resulting in an undesired outcome when pivoting the data

voter_data[(3536),] <- voter_data[(3536),] %>%
    mutate(candidatevotes = 1677928 + 78)
voter_data[(3537),] <- voter_data[(3537),] %>%
    mutate(candidatevotes = 943169 + 259)

# Manually rename more conflicting cells
# In Minnesota, for some years, the democrat party was entered as "democratic". Our algorithm didn't pick up the string and resulted in NA's

voter_data[(2211),] <- voter_data[(2211),] %>%
    mutate(party = "democrat")
voter_data[(2560),] <- voter_data[(2560),] %>%
    mutate(party = "democrat")
voter_data[(3226),] <- voter_data[(3226),] %>%
    mutate(party = "democrat")

# Eliminate conflicting rows

voter_data <- voter_data[-c(2545,3415,3543,3544),]

```

Once the conflicts have been fixed, we're ready to build a data frame with voting data that will be friendly with our model.

```{r build_clean_voter_data}
clean_votes <- voter_data %>%
  # Choose only the two main parties
  filter (party %in% c("democrat", "republican")) %>% 
  # Select the only variables we're interested in. Year, state, party, and how many votes the party received
  select(year, state, party, candidatevotes) %>% 
  pivot_wider(names_from = party, values_from = candidatevotes) %>%
  rename(democrat_votes = democrat, republican_votes = republican) %>%
  # Create a new variable that tells us who won the election in that state
  mutate(office = ifelse(democrat_votes > republican_votes, "democrat", "republican")) %>%
  # Filter so voting data is consistent with the rest of our data
  filter(year >= 1976)

glimpse(clean_votes)
```

```{r model_production}

### Build production data frame
model_production <- beer_states %>%
  # Chose only years previous to 2015
  filter(year <= 2015) %>%
  # Clean names
  mutate(state = str_to_title (state)) %>%
  pivot_wider(names_from = type, values_from = barrels)


```





```{r model_dataframe, message = FALSE}
 
  

### Build model data frame by joining production and votes

model_data_production <- full_join(model_production, clean_votes) %>%
  # Eliminate regions and DC, which doesn't have voter representation.
  filter(!(state %in% c("District Of Columbia", "District of Columbia", "Us Total", "South Region", "Northeast Region", "Midwest Region", "West Region"))) %>%
  select(c("year", "state","On Premises", "Bottles and Cans","Kegs and Barrels", "office")) %>%
  filter(year >= 2008) %>%
  arrange(state, year)

glimpse(model_data)

### Split data into train and test

model_data_test <- model_data %>%
  filter(year == 2015)

model_data_train <- model_data %>%
  filter(year < 2015)

production_test <- model_data_production %>%
  filter(year == 2016)

production_train <- model_data_production %>%
  filter(year != 2016)

for (n in 1:nrow(production_train)) {
  if ((n-1)%%4 == 0) {
  production_train$office[n] -> production_train$office[n+3]
  production_train$office[n] -> production_train$office[n+2]
  production_train$office[n] -> production_train$office[n+1]
  }
}

# Apply a simple "for" loop that will "fill in" what party that state voted for in the last four years.

for (n in 1:nrow(model_data_train)) {
  if ((n-1)%%4 == 0) {
  model_data_train$office[n] -> model_data_train$office[n+3]
  model_data_train$office[n] -> model_data_train$office[n+2]
  model_data_train$office[n] -> model_data_train$office[n+1]
  }
}

# Drop out the year 1976. It was only being used to fill in "office" values.

model_data_train <- model_data_train %>%
  filter(year != 1976)

# Since we want to predict "office", we have to convert it into a factor.

model_data_train["office"] <- lapply(model_data_train["office"] , factor)

glimpse(model_data_train)
  
```

### Modelling

We will fit a logistic regression model using our training data 

```{r fitting_model}

model_election <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(office ~ ., data = model_data_train, family = "binomial")

tidy(model_election)
```
We will now predict each states winner in the 2016 election. For this we will create a data frame where we compare our predictions to the real values.

```{r testing_model}

election_pred <- predict(model_election, model_data_test, type = "prob") %>%
  bind_cols(model_data_test %>% select(office, state)) %>%
    mutate(predicted = ifelse(.pred_democrat > .pred_republican, "democrat", "republican")) %>%
  select(.pred_democrat, .pred_republican, state, office, predicted) %>%
  print(n = 10)

```


```{r plotting_compared_results}

partisan_colors <- c("democrat" = "blue3", "republican" = "red3", "null" = "black")

plot_usmap(data = (election_pred %>% filter(predicted == office)), regions = "states", values = "office", color = "black") +
  labs( title = "Election winners predicted correctly",
        subtitle = "by state",
        caption = "States coloured white were predicted incorrectly",
        fill = "Party") +
  theme(
    legend.position = "right") +
  scale_fill_manual(values = partisan_colors)


election_pred %>%
  count(office == predicted)

```
```{r write_csv}

write_csv(election_pred, "/cloud/project/data/clean_presentation_data/election_pred.csv")
write_csv(model_data_train, "/cloud/project/data/clean_presentation_data/model_data_train.csv")
write_csv(clean_votes, "/cloud/project/data/clean_presentation_data/clean_votes.csv")
write_csv(model_production, "/cloud/project/data/clean_presentation_data/model_production.csv")

```

```{r find-winning-party}
election_out <- full_join(election_pred, electoral_votes) %>%
  filter(state != "District Of Columbia")

outcome <- election_out %>% 
  group_by(predicted) %>%
  summarise(total_votes = sum (votes))

winner <- outcome %>%
  mutate(max = max(total_votes)) %>%
  filter(total_votes == max)

winner$predicted
```


